---
title: "Penguins ANOVA"
author: "Bridget Gibbons"
date: "11/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##1. Load packages and read-in penguins.csv data

```{r, include = FALSE}
library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")
```

Our question:

Is there a significant difference in mean flipper length for femail penguins by species (Chinstrap, Gentoo, and Adelie)?

##2. Do some exploring

New functions: with + table for easy counting over multiple factors

Update species name using case_when

```{r}

counts <- with(penguins, table(species, sex))

penguins2 <- penguins %>% 
  mutate(common_name =
           case_when(
             species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
             species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
             species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
           )
         ) %>% 
  select(-species) %>% 
  filter(sex == "FEMALE")

```

- Exploratory graphs
- Summary statistics
- Levene's test for equal variances
- one-way ANOVA
- post-hoc testing

```{r}

hists <- ggplot(penguins2, aes(x = flipper_length)) +
  geom_histogram(aes(fill = common_name)) + 
  facet_wrap(~ common_name)

hists

qqs <- ggplot(penguins2, aes(sample = flipper_length)) +
  geom_qq() +
  facet_wrap(~ common_name)
  
qqs


# Levene's test for equal variances (>2 groups... if you have two groups, use the F-test for equal variances)

# We'll use the function leveneTest from the 'car' package

# H0: Variances are equal
# HA: Variances are unequal

penguin_levene <- leveneTest(flipper_length ~ common_name, data = penguins2)
penguin_levene

# We retain the null hypothesis of equal variances (p = 0.06)

penguin_variances <- penguins2 %>% 
  group_by(common_name) %>% 
  summarize(
    variance = var(flipper_length)
  )

penguin_variances

#Even though we got significant result for Levene's test, if your largest variance is less than 4x the smallest variance, can still use ANOVA
```

##3. One-way ANOVA

Single factor: common_name (species)
Number of levels in factor: 2 (Chinstrap, Gentoo, Adelie)
Random variable: flipper_length (mm)

Is there a significant difference in mean flipper length (mm) for female Chinstrap, Gentoo, and Adelie penguins?

```{r}
# H0: Mean flipper lengths across all species are equal
# HA: At least two means differ significantly

penguin_aov <- aov(flipper_length ~ common_name, data = penguins2)
summary(penguin_aov)

# If we take sampels for these three groups and they actually have the same mean flipper length, there is nearly a 0% chance that we would have gotten the difference in one or more means that we got, by random chance.

# At least two samples were taken from populations with different means. Which ones are different? All three are different from eachother? Or something else?

# Post-hoc testing using Tukey's HSD

penguin_ph <- TukeyHSD(penguin_aov)
penguin_ph

```

Mean flipper lengths for Adelie, Gentoo, and Chinstrap penguins are all significantly different (F(df) = statistic, p <0.001, $\alpha$ = 0.05 with post-hoc Tukey's HSD, $\alpha = 0.05$.

Ex. Of the five locations studied, two locations were found to differ significantly (include ANOVA and the Tukey's info)